---
title: "Penguins Lab"
author: "Daphne Virlar"
date: "11/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Load packages and read in penguin data

Packages: tidyverse, car, onewaytests
```{r}

library(tidyverse)
library(car)
library(onewaytests)

penguins <- read_csv("penguins.csv")

```


*Our question: Is there a significant difference in mean flipper lengths for female Chinstrap, Gentoo, and Adelie penguins near Palmer Archipelago?*

Maybe I can use a one-way ANOVA...
-Independent observations
-Equal Variance (levene's test)
-Normality (histograms / qqplot)


```{r, wrangle}

counts <- with(penguins, table(species, sex))
View(counts)

#Update the names of colummns to make data more manageable
penguins2 <- penguins %>% 
  mutate(
    common_name = 
      case_when(
        species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
        species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
        species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
      )
  ) %>% 
  select(-species) %>% 
  filter(sex == "FEMALE")



```


Exploratory graphs 
```{r}


hists <- ggplot(penguins2, aes(x = flipper_length))+
  geom_histogram(aes(fill = common_name))+
  facet_wrap(~common_name, scales = "free")

hists


peng_quant <- ggplot(penguins2, aes(sample = flipper_length))+
  geom_qq(aes(color = common_name))+
  facet_wrap(~ common_name)

peng_quant

# Based on histograms and qq-plots, observations for each group appear normall distributed. Also, n > 30 for each group, so means will be normally distributed by CLT.

```


Evaluate variances and perform a Levene's test for equal variances
```{r}
#If the largest sample variance is < 4x greater than the smallest sample variance, then these are "close enough" to consider variances equal

variances <- penguins2 %>% 
  group_by(common_name) %>% 
  summarize(
    variance = var(flipper_length)
  )


variances
  # the largest variance (33.1) is definitely not more than 4x greater than the smallest variance (15.2)


# formal hypothesis test
# Levene's Test
      # leveneTest function in the 'car' package

levene_peng <- leveneTest(flipper_length ~ common_name, data = penguins2)
levene_peng


# H0: Variances are equal
# HA: Variances are NOT equal

# There is no significant difference in variances across species.

```


One-Way ANOVA for penguin flipper lengths by species (common-name)

```{r}

# Random Variable: Flipper length
# Number of Factors: 1
# No. of Levels: 3
# Factor: Species
# Levels: Chinstrap, Gentoo, and Adelie


# H0: Means across all groups are equal
# HA: At least TWO means are differe significantly


penguin_anova <- aov(flipper_length ~ common_name, data = penguins2)
summary(penguin_anova)


# Post-hoc testing by Tukey's HSD
  # HO: Pairwise means are the same
  # HA: Pairwise means are NOT the same
penguin_ph <- TukeyHSD(penguin_anova)
penguin_ph


# aov.test function in the 'onewaytests' package DOES allow for inline referencing.
```

Mean flipper lengths of female penguins differed significantly for all species studied (one-way anova; F(DF) = statistic, p < 0.001, $\alpha$ = 0.05 with post-hoc Tukey's HSD, $\alpha$ = 0.05).